My title

### 2024-10-29 paper/paper.qmd
added TODOs

---
title: "Forecasting the United States 2024 Presidential Election"
subtitle: "Kamala Harris wins 60th presidency with 265 Electoral College Votes"
author: 
  - Robert Ford
  - Michelle Ji
  - Cher Ning
thanks: "Code and data are available at: https://github.com/Ford-Robert/us-presidential-election."
date: today
date-format: long
abstract: "This report analyzes compiled polling surveys for the 2024 U.S. presidential election using FiveThirtyEight data [@538data] and the 'pool of polls' method, which aggregates multiple polls to create a stable estimate of public opinion and reduce bias. From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system. The models forecast Kamala Harris as the winner of election, winning 265 electoral college votes versus the 172 votes Donald Trump acquires. Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process. Additionally, it helps predict potential policy shifts and holds political institutions accountable, contributing to a more transparent and informed democracy."
format: pdf
header-includes:
  - \usepackage{geometry}
  - \geometry{margin=0.75in} 
number-sections: true
bibliography: references.bib
---

@@

Every four years, the United States presidential election occurs and is one of the most significant political events internationally. The complex process of the election involves both the popular and electoral college system. The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege]. After the popular vote is accounted for, the candidate who receives the majority of the electoral votes from all states wins the presidential election. The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege].

Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs. For example, regarding technology, both parties are concerned with the rapid progression of artificial intelligence’s capabilities but differ in the response and regulation of it. Republicans believe that the current practices are unfair and disproportionately target their voices, whereas democrats generally are in favor of tighter regulation [@keyissues]. Another key issue important to voters and the candidates is abortion. As former president, he supported restrictions on abortion access and advocated for policies that align with the conservative movement’s opposition to abortion rights [@abortionstance]. On the other hand, Harris has been a key supporter in the pro-choice movement and has been an advocate for protecting and expanding access to abortion services, as well as female autonomy [@harrisviews]. The outcome of this election will have profound consequences for various legislative policies and the future direction of the United States.

@@

The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region. Pollster Name indicates which pollster the individual poll came from and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time to adjust for any bias. Method is the way the poll was deployed, including but not limited to live phone, online panel, and text-to-web. State indicates what state the poll was conducted and sample size is the size of the poll. The pollscore is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number. The election date and days to election variables indicate how far out the poll was conducted with respect with the election date of November 5, 2024. We also reorganized the dataset to include a poll region, which divides the US into geographical regions such as rust belt, northeast, etc. The table in @sec-app-table details the first few observations of the cleaned dataset, which we used to build our model.

@@

```{r}
#| fig-cap: Top 25 pollsters in overall polling data
#| echo: false
#| warning: false
#| message: false

@@

@@

@@

## Analysis into a Pollster's Methodology

One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article). Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants’ responses on 51 questions through either live phonecall interviews or an online survey. The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article). This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.

The key finding from this poll is that 48% of participants favoured Harris while 50% favoured Trump, with the lead being consistent in the larger sample of registered voters and smaller subsample of likely voters. Despite this, responses also indicate that Harris is narrowly leading in seven swing states, meaning that Democrats could potentially win by electoral college while losing the popular vote (cite Fox article). This report also notes that current results are Trump’s highest approval ratings since Biden dropped out of the race, while support for Harris is at its lowest.

@@

On the other hand, one potential weakness of this poll is that it was sponsored by Fox News, a news site that has historically been Republican-leaning. Though from the reported methodology alone, no evident biasing of the pollster can be observed, this potential source of biased reporting must be noted as it can affect the type of analyses that occurred and which key findings are the focus of news reports. As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article). This is a critical area to pay attention to as it is an option offered on every question, selected by up to 4% of participants on key questions such as “If the presidential election were today, how would you vote?” (cite Fox article).

## Idealized Methodology and Survey

## Table Detailing First Few Observations of Dataset {#sec-app-table}

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Robert Ford
  - Michelle Ji
  - Cher Ning
thanks: "Code and data are available at: https://github.com/Ford-Robert/us-presidential-election."
date: today
date-format: long
abstract: "Compiled polling surveys regarding the 2024 United States presidential election were analyzed in this report based on data from FiveThirtyEight [@538data]. From the analysis, bayesian and generalized linear models were built to forecast the outcome of the presidential election based on popular vote and the electoral college system. The models forecast Kamala Harris as the winner of the US presidency. Election forecasting is crucial for informing voters, guiding campaign strategies, and engaging the public in the electoral process. Additionally, it helps predict potential policy shifts and holds political institutions accountable, contributing to a more transparent and informed democracy."
format: pdf
header-includes:
  - \usepackage{geometry}
  - \geometry{margin=0.75in} 
number-sections: true
bibliography: references.bib
---

@@

Every four years, the United States presidential election occurs and is one of the most significant political events internationally. The complex process of the election involves both the popular and electoral college system, in which voters cast ballots on election day in each state. The candidate that receives the popular vote by the voters wins the electoral votes for that state [@electoralcollege]. After the popular vote is accounted for, the candidate who receives the majority of the electoral votes from all states wins the presidential election. The number of electoral votes varies by state and is proportional to each state’s respective population [@electoralcollege]. Specifically for the 2024 presidential election, the top candidates include Donald Trump and Kamala Harris, both with differing opinions on key topics such as immigration, technology, abortion, and transatlantic affairs. For example, regarding technology, both parties are concerned with the rapid progression of artificial intelligence’s capabilities but differ in the response and regulation of it. Republicans believe that the current practices are unfair and disproportionately target their voices, whereas democrats generally are in favor of tighter regulation [@keyissues]. Another key issue important to voters and the candidates is abortion. As former president, he supported restrictions on abortion access and advocated for policies that align with the conservative movement’s opposition to abortion rights [@abortionstance]. On the other hand, Harris has been a key supporter in the pro-choice movement and has been an advocate for protecting and expanding access to abortion services, as well as female autonomy [@harrisviews]. The outcome of this election will have profound consequences for various legislative policies and the future direction of the United States.

@@

The variables used in the dataset include pollster name, pollster weight, method, state, sample size, pollscore, collection/end date, transparency score, candidate, support percentage, election date, days to election, initial weight, weight, and poll region. Pollster Name indicates which pollster the individual poll came from, which varies from individual companies, such as YouGov, to research institutions, such as Quinnipiac University and the weight is assigned by FiveThirtyEight depending on sample size and if they have multiple polls out in a short time. This is to adjust for any bias. Method is the way the poll was deployed, including but not limited to live phone, online panel, and text-to-web. State indicates which state the poll was conducted. State indicates what state the poll was conducted and sample size is the size of the poll. The pollscore is the rating FiveThirtyEight gives each indidivual poll, on factors such on ethical and methodological elements.The candidate variable indicates the candidate a certain percentage of people in the poll voted for and "support percentage” represents that specific number. The election date and days to election variables indicate how far out the poll was conducted with respect with the election date of November 5, 2024. We also reorganized the dataset to include a poll region, which divides the US into geographical regions such as rust belt, northeast, etc.

The table details the first few observations of the cleaned dataset, which we used to build our model.

@@

```{r}
#| label: fig-1
#| fig-cap: Top 25 pollsters in overall polling data
#| echo: false
#| warning: false
#| message: false

@@

TODO NEED TO JUSTIFY MY PRIORS MORE

@@

FOR IDEALIZED SURVEY
!TODO like the other appendix, need to talk about depth about the techniques. Need to create a google form that is the actual survey. Every decision made about the idealized survey must be justified with literature. Every detail of the survey must be accounted for. Need to write about how the survey was created, and identify and justify all of the tradeoff made. TODO!

@@

!TODO need to add details about the techniques used by the pollster. For example, if stratified sampling was used, dive into what is stratified sampling. What is it good for why is works. Reference a academic paper about stratified sampling. TODO!

One particular poll within our data sample is one conducted between October 11-14 by Beacon Research and Shaw & Company Research (cite Fox article). Their analysis found 48% of participants favoured Harris while 50% favoured Trump. Targeting the sample frame of registered American voters, this poll collected data on 1,110 participants through either live phonecall interviews or an online survey. The sample was recruited by applying the probability proportionate to size method on the nationwide voter file of registered voters’ phone numbers, meaning that the contacted individuals would be proportionally representative of the number of voters per state (cite Fox article). This ensures that the voices represented by poll results would match the ones involved with the real election as much as possible.

@@

	On the other hand, one potential weakness of this poll is that it was sponsored by Fox News, a news site that has historically been Republican-leaning. Though from the reported methodology alone, no evident biasing of the pollster can be observed, this potential source of biased reporting must be noted as it can affect the type of analyses that occurred and which key findings are the focus of news reports. As well, another potential issue is the lack of indication regarding how non-responses of “Don’t know” were handled during analysis (cite Fox article).
